---
title: "algorithms"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the required libraries

```{r}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(tidyquant)
library(magrittr)
```

# Create utility functions

```{r}
get_cols_idxs <- function(df, col_names) {
  return(which(names(df) %in% col_names))
}
```

# Load the and treat the data

```{r}
algorithms <- read.csv(file="uniform_start.csv")
algorithms$file <- sub("^(\\w+)_(\\d+)", "\\2", algorithms$file)
algorithms$file <- as.numeric(algorithms$file)
```
# Show basic statistics for each algorithm for each delta

```{r}
algorithms %>% 
  group_by(algorithm, file) %>% 
  summarize(
    avg_rscore=mean(Rscore_Algorithm_Capacity), 
    sd_rscore=sd(Rscore_Algorithm_Capacity),
  ) %>%
  arrange(-file, avg_rscore)
```

# Plot relevant data

## Rscore min difference

For each iteration, the minimum Rscore is computed, and for each algorithm, the value for that iteration is the difference of the algorithms Rscore and the minimum value. 

This implies that the Rscore min difference will never be smaller than 0. 

Given the table with the computed differences, the average is taken for each tested delta producing the final graph of the algorithms performance for the delta test.

### Procedure

First we turn each algorithm into their own column so we can then use it to compute the min difference
```{r}
wider_rscore <- algorithms %>%
  summarize(file, iteration, algorithm, Rscore_Algorithm_Capacity) %>%
  pivot_wider(names_from=algorithm, values_from=Rscore_Algorithm_Capacity)

wider_rscore$min_Rscore <- apply(wider_rscore[, -get_cols_idxs(wider_rscore, c("file", "iteration"))], 1, FUN=min, na.rm=TRUE)
wider_rscore[, -get_cols_idxs(wider_rscore, c("file", "iteration"))] %<>% 
  mutate_at(vars(-min_Rscore), list(~ (. - min_Rscore)))
```

Having computed the difference, the algorithms are now pivoted longer, to allow for tidyr grouping operations.

```{r}
longer_rscore <- wider_rscore %>% pivot_longer(-c("file", "iteration", "min_Rscore"), names_to="algorithm", values_to="diff")
grouped_rscore <- longer_rscore %>% group_by(file, algorithm) %>% summarize(avg_diff=mean(diff)) # %>% filter(algorithm %in% c("bfd", "mbfp", "mbf", "mwf", "mwfp"))

ggplot(grouped_rscore, aes(x=file, y=avg_diff, color=algorithm)) + 
  geom_point() + 
  geom_line(data=grouped_rscore, aes(group=algorithm)) + 
  labs(title="RSCORE diff")
```


## Difference to minimum number of consumers

### Procedure

```{r}
wider_nconsumers <- algorithms %>% 
  summarize(file, iteration, algorithm, Number.of.Consumers) %>%
  pivot_wider(names_from=algorithm, values_from=Number.of.Consumers)

wider_nconsumers$min_consumers <- apply(wider_nconsumers[,-get_cols_idxs(wider_nconsumers, c("file", "iteration"))], 1, FUN=min, na.rm=TRUE)
wider_nconsumers[, -get_cols_idxs(wider_nconsumers, c("file", "iteration"))] %<>% 
  mutate_at(vars(-min_consumers), list(~ (. - min_consumers)/min_consumers))
```

```{r}
longer_nconsumers <- wider_nconsumers %>% pivot_longer(!c("file", "iteration", "min_consumers"), names_to="algorithm", values_to="diff") 
group_nconsumers <- longer_nconsumers %>% group_by(file, algorithm) %>% summarize(avg_diff_to_min=mean(diff)) 

ggplot(group_nconsumers, aes(x=file, y=avg_diff_to_min, color=algorithm)) + 
  geom_point() + 
  geom_line(data=group_nconsumers, aes(group=algorithm)) +
  labs(
    y="Average Relative Deviation from minimum number of consumers", 
    x="Delta"
  )
```

```{r}
filtered_nconsumers <- group_nconsumers %>% 
  filter(algorithm %in% c("mwf", "mbf", "mbfp", "mwfp", "bfd")) #, "mbfp", "mbf", "mwf", "mwfp"))

ggplot(filtered_nconsumers, aes(x=file, y=avg_diff_to_min, color=algorithm)) + 
  geom_point() + 
  geom_line(data=filtered_nconsumers, aes(group=algorithm)) +
  labs(
    y="Average Relative Deviation from minimum number of consumers", 
    x="Delta"
  )

```

## Average Rscore

```{r}
rscore_avg <- algorithms %>% 
  group_by(file, algorithm) %>% 
  summarize(avg_rscore=mean(Rscore_Algorithm_Capacity))

ggplot(rscore_avg, aes(x=file, y=avg_rscore, color=algorithm)) + 
  geom_point() + geom_line(data=rscore_avg, aes(group=algorithm)) + 
  labs(
    y="Average Rscore", 
    x="Delta"
  )
```

